---
title: "Apertium Translation with polyglotr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Apertium Translation with polyglotr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(polyglotr)
```

# Introduction to Apertium Translation

This vignette introduces the Apertium translation functionality in the `polyglotr` package. Apertium is an open-source machine translation platform that provides free translation services without requiring API keys or authentication.

## What is Apertium?

[Apertium](https://apertium.org/) is a free, open-source machine translation platform that:

- Provides rule-based machine translation
- Requires no authentication or API keys
- Offers translation between many language pairs
- Has a public API available at `https://apertium.org/apy/`
- Aligns perfectly with polyglotr's philosophy of providing access to free translation services

## Basic Usage

The `translate_apertium()` function provides a simple interface to translate text using the Apertium API:

```{r basic-usage}
# Basic translation from English to Spanish
result <- translate_apertium(
  text = "Hello, world!", 
  target_language = "es", 
  source_language = "en"
)
print(result)

# Translation from Spanish to English
result <- translate_apertium(
  text = "Hola, mundo!", 
  target_language = "en", 
  source_language = "es"
)
print(result)
```

## Checking Available Language Pairs

Before translating, you can check which language pairs are supported by making a request to the `/listPairs` endpoint:

```{r list-pairs, eval=FALSE}
# Example of how to check available language pairs (manual approach)
# This would typically be done outside of R or with a custom function
library(httr)
library(jsonlite)

response <- GET("https://apertium.org/apy/listPairs")
pairs <- content(response)

# The response contains an array of objects like:
# {"sourceLanguage":"en","targetLanguage":"es"}
head(pairs, 10)
```

## Translating Multiple Texts

You can translate multiple pieces of text by using vectorization or loops:

```{r multiple-texts}
# Vector of texts to translate
texts <- c(
  "Good morning", 
  "How are you?", 
  "Thank you very much",
  "See you later"
)

# Translate each text
translations <- sapply(texts, function(text) {
  translate_apertium(
    text = text,
    target_language = "es",
    source_language = "en"
  )
})

# Display results
data.frame(
  English = texts,
  Spanish = translations,
  stringsAsFactors = FALSE
)
```

## Custom Host Configuration

The function allows you to specify a custom host if you're using a different Apertium instance:

```{r custom-host}
# Using default host (https://apertium.org/apy)
translate_apertium("Hello", "es", "en")

# Using a custom host (if available)
translate_apertium(
  text = "Hello", 
  target_language = "es", 
  source_language = "en",
  host = "https://your-custom-apertium-server.com/apy"
)
```

## Error Handling

The function includes basic error handling for common issues:

```{r error-handling}
# Example of handling potential errors
tryCatch({
  result <- translate_apertium("Hello", "invalid-lang", "en")
  print(result)
}, error = function(e) {
  cat("Translation error:", e$message, "\n")
})
```

## Comparison with Other Services

Apertium offers several advantages within the polyglotr ecosystem:

1. **No API Keys**: Unlike Google Translate or Microsoft Translator, Apertium requires no authentication
2. **Open Source**: The entire platform is open source and transparent
3. **Free**: No usage limits or costs
4. **Privacy**: Requests are not tracked for commercial purposes
5. **Rule-based**: Uses linguistic rules rather than statistical methods

```{r comparison}
# Compare translations across services
text <- "The weather is beautiful today"

# Apertium translation
apertium_result <- translate_apertium(text, "es", "en")

# MyMemory translation (also free, but different approach)
mymemory_result <- mymemory_translate(text, "es", "en")

data.frame(
  Service = c("Apertium", "MyMemory"),
  Translation = c(apertium_result, mymemory_result),
  stringsAsFactors = FALSE
)
```

## Best Practices

When using Apertium translation:

1. **Check Language Support**: Verify that your desired language pair is supported
2. **Handle Network Errors**: Always wrap calls in error handling for network issues
3. **Batch Processing**: For multiple translations, consider adding delays to be respectful to the service
4. **Text Preprocessing**: Clean your text of special characters that might cause issues

```{r best-practices}
# Example of robust translation function
safe_translate_apertium <- function(text, target_lang, source_lang, max_retries = 3) {
  for (attempt in 1:max_retries) {
    tryCatch({
      result <- translate_apertium(text, target_lang, source_lang)
      return(result)
    }, error = function(e) {
      if (attempt == max_retries) {
        stop("Translation failed after ", max_retries, " attempts: ", e$message)
      }
      Sys.sleep(1)  # Wait before retry
    })
  }
}

# Usage
result <- safe_translate_apertium("Hello world", "es", "en")
print(result)
```

## Conclusion

The `translate_apertium()` function provides a simple, free, and open-source solution for machine translation within the polyglotr package. It's particularly useful for users who want to avoid API keys and usage limits while still accessing quality machine translation services.

For more information about Apertium, visit [https://apertium.org/](https://apertium.org/) and explore their documentation at [https://apertium.org/apy/](https://apertium.org/apy/).