---
title: "Translation of Movie Reviews"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Translation of Movie Reviews}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
In this vignette, we will demonstrate how to use the `polyglotr` package in combination with other packages like text2vec, purrr, and dplyr to translate movie reviews into French. 

# Installation

To get started, you need to install the required packages. If you haven't installed the Polyglotr package and the other dependencies, you can use the following code to install them:

```{r setup}
library(polyglotr)
library(purrr)
library(dplyr)
library(text2vec)
```

# Translating Movie Reviews

To demonstrate the translation of movie reviews, we will use the movie_review dataset from the text2vec package. In the code below, we load the necessary packages and create a subset of the `movie_review` dataset with the first 10 rows, which will be used for the translation.

```{r reviews}
df <- head(movie_review, 10)
glimpse(df)

## show reviews
df$review
```

Now, let's translate the review column to French using the `polyglotr` package:

```{r translated}
# Translate the review column to French
translated_reviews <- df %>%
  dplyr::mutate(french_review = purrr::map_chr(review, ~ google_translate(.x, target_language = "fr", source_language = "en")))
```


In the code snippet above, we utilize the `mutate` function from `dplyr` in conjunction with `purrr::map_chr` to apply the google_translate function from the package to each element in the review column. The translated reviews are stored in a new column named "french_review".

# Results

Let's examine the translated movie reviews:

```{r results}
glimpse(translated_reviews)

translated_reviews$french_review
```


The output will display the first few rows of the translated movie reviews, including the original review column and the corresponding French translation in the "french_review" column.



